---
title: "p8105_hw1_yz4982"
author: "Yuanyuan Zhang"
output:
  html_document:
    toc: true
    toc_float: true
---

Promblem 1
------------------------------------------------------------
Use the early_january_weather dataset from the moderndive package.
```{r load-weather}
# load the moderndive package
library(moderndive)

# Load the early_january_weather dataset
data("early_january_weather")

# Show the first 6 rows 
head(early_january_weather)

# Show dataset structure with variable types
dplyr::glimpse(early_january_weather)

```

The variables in this dataset are:  
`r paste(names(early_january_weather), collapse = ", ")`.  

Among these, some key variables are:  
**origin**: indicates the weather station code, with observed values: 
`r paste(unique(early_january_weather$origin), collapse = ", ")`.  
**month**: records the month of the observation, with unique value(s): 
`r paste(unique(early_january_weather$month), collapse = ", ")`.  
**day**: records the calendar day of the month, ranging from 
`r min(early_january_weather$day)` to `r max(early_january_weather$day)`.    
**temp**: temperature in °F, with a mean of 
  `r round(mean(early_january_weather$temp, na.rm = TRUE), 1)`.  
**humid**: relative humidity (0–1 scale), ranging from 
  `r round(min(early_january_weather$humid, na.rm = TRUE), 2)` to 
  `r round(max(early_january_weather$humid, na.rm = TRUE), 2)`.  
**wind_speed**: average wind speed in mph, with a mean of 
  `r round(mean(early_january_weather$wind_speed, na.rm = TRUE), 1)`.  
**pressure**: atmospheric pressure in millibars, with a mean of 
  `r round(mean(early_january_weather$pressure, na.rm = TRUE), 1)`.  
**time_hour**: hourly date-time observations, ranging from 
  `r min(early_january_weather$time_hour)` to 
  `r max(early_january_weather$time_hour)`. 

The dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns. 

The average temperature is `r round(mean(early_january_weather$temp, na.rm = TRUE), 1)` °F.


Scatterplot of temperature vs time
```{r p1-plot, fig.width=8, fig.height=4}
library(ggplot2)

p1 <- ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point(alpha = 0.7) +
  labs(
    title = "Hourly Temperature in Early January",
    x = "Time (hourly)",
    y = "Temperature (°F)",
    color = "Humidity"
  ) +
  theme_minimal()

p1
```
